Arbitrary Precision and Arbitrary $n$-grams

### 2011-02-28 paper/paper.tex
Changes, fixes, etc.

\documentclass{article}
\newcommand{\osn}{\oldstylenums}
\title{TITLE HERE}
\author{Alec Story and Thomas Levine\\\small{avs\osn{38} and tkl\osn{22}}}

@@

	Each subsection contains two examples from the Shakespeare corpus and two
	from War and Peace.  These corpora were chosen because they were determined
	at rough glance to be the easiest to parse and had the fewest problematic

@@

@@

		One of the two extensions we chose was to implement the probability and
		perplexity calculations using arbitrary precision arithmetic (we also
		implemented it using logarithmic arithmetic).  This can lead to slower
		processing, but in our profiling, most of the time was spent in
		dictionary manipulation rather than mathematics, and guarantees that no
		information is lost.

		We used python's +gmpy+ package to represent the probabilities as
		arbitrary rational numbers (python's integers are already
		arbitrary-precision), and +mpmath+ to perform the final exponentiation
		to calculate the perplexity (the $n$-th root).

@@

		Because the computations were general enough, we were able to easily
		abstract the code to produce count, smoothed count, and probability
		tables for arbitrary $n$-grams.  We also compute all of the smaller
		$n$-grams, which are required for generating or analyzing the beginning
		of a sequence of words, and for other computations.
\end{document}

>>>>

\documentclass{article}
\newcommand{\osn}{\oldstylenums}
\title{Arbitrary Precision and Arbitrary $n$-grams}
\author{Alec Story and Thomas Levine\\\small{avs\osn{38} and tkl\osn{22}}}

@@

	Each subsection contains two examples from the Shakespeare corpus and two
	from War and Peace.  These corpora were chosen because they were determined
	at rough glance to be the easiest to parse and had the fewest problematic.

@@

		Trigrams lead to more comprehensible sentences, which is to be expected,
		and bigrams are barely useful.  The unigrams also seemed to produce
		longer sentences; it's possible that trigrams and bigrams lead to a
		progression from words that are likely at the beginning of sentences to
		words that are likely at the end, and then to end the sentence, despite
		sentence ends being fairly infrequent.  Additionally, \textit{War and
		Peace} seems to produce longer sentences (particularly when observed
		over more iterations than can be reproduced comfortably here), probably
		because the sentences in the original are longer, so sentence ends are
		rarer.

		Our $n$-gram generator only treated sentence endings by adding a single
		symbol, `\#', and did not break the input into individual sentences, nor
		do we break the perplexity input into sentences in the same manner.

@@

		One of the two extensions we chose was to implement the probability and
		perplexity calculations using arbitrary precision arithmetic (we also
		implemented it using logarithmic arithmetic).  This can lead to slower
		processing, but in our profiling, most of the time was spent in
		dictionary manipulation rather than mathematics.  This approach
		guarantees that there is no data loss until we move into processing
		floats during exponentiation.

		We used python's +gmpy+ package to represent the probabilities as
		arbitrary rational numbers (python's integers are already
		arbitrary-precision), and +mpmath+ to perform the final exponentiation
		to calculate the perplexity (the $n$-th root).  The libraries could only
		compute non-integer powers of integers or floats, and integer powers
		were rounded to the nearest integer, and therefore, inaccurate, so
		(very high precision) floats provided the best results.  This final
		stage of computation of perplexity was the only point where imprecise
		mathematics was used.

@@

		Because the computations were general enough, we were able to easily
		abstract the code to produce count, smoothed count, and probability
		tables for arbitrary $n$-grams.  We also compute all of the smaller
		$n$-grams, which are required for generating or analyzing the beginning
		of a sequence of words, and for other computations.

		Large $n$-grams (particularly $n > 4$) tend to result in whole
		Shakespeare sentence fragments.  For example, ``O , my offence is rank ,
		it smells to heaven ; It hath the primal eldest curse upon't , A
		brother's murther !'' was generated by a hexagram, and is a direct quote
		from Hamlet, act \osn{3} scene \osn{3}.  This is a well-documented
		phenomenon, and is due to the fact that it's quite possible that the
		preceding $n-1$-gram appears only once in the entire corpus, so the
		sentence generator has no choice but to choose the next word in the
		quotation.

\end{document}
